---
title: "hw2_ryw2109"
author: "Rita Wang"
date: "2024-10-02"
output: github_document
---

### Question 0
```{r libraries, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(readxl) # Needed to import excel files
```

```{r importing_data, echo = FALSE, message = FALSE, warning = FALSE}
# Question 1 dataset
transit_df = read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names()
transit_df

# Question 2 dataset
trash_df = read_excel("./data/202409 Trash Wheel Collection Data.xlsx")%>% 
  janitor::clean_names()
trash_df

# Question 3 dataset
bakers_df = read_csv("./data/gbb_datasets/bakers.csv")%>% 
  janitor::clean_names()
bakers_df

bakes_df = read_csv("./data/gbb_datasets/bakes.csv")%>% 
  janitor::clean_names()
bakes_df

results_df = read_csv("./data/gbb_datasets/results.csv")%>% 
  janitor::clean_names()
results_df
```

### Question 1
```{r}
transit_df = read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%  #import data
  janitor::clean_names() %>%  #clean column name
  select(line, station_name, station_latitude, station_longitude, route1 : route11, entry, vending, entrance_type, ada) %>% #retain the columns desired
  mutate(entry = 
    ifelse(entry == "Yes", TRUE, FALSE
    )
          )
transit_df


```
Write a short paragraph about this dataset – explain briefly what variables the dataset contains, describe your data cleaning steps so far, and give the dimension (rows x columns) of the resulting dataset. Are these data tidy?

How many distinct stations are there? Note that stations are identified both by name and by line (e.g. 125th St 8th Avenue; 125st Broadway; 125st Lenox); the distinct function may be useful here.

How many stations are ADA compliant?

What proportion of station entrances / exits without vending allow entrance?

Reformat data so that route number and route name are distinct variables.

How many distinct stations serve the A train? Of the stations that serve the A train, how many are ADA compliant?

### Question 2
```{r}



```