---
title: "hw2_ryw2109"
author: "Rita Wang"
date: "2024-10-02"
output: github_document
---

### Question 0
```{r libraries, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(readxl) # Needed to import excel files
```

### Question 1
```{r Q1}
transit_df = read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%  #import data
  janitor::clean_names() %>%  #clean column name
  select(line, station_name, station_latitude, station_longitude, route1 : route11, entry, vending, entrance_type, ada) %>% #retain the columns desired
  mutate(entry = 
    ifelse(entry == "Yes", TRUE, FALSE
    ) #changes the Yes/No of the Entry column into logical values
          )
transit_df

# total distinct stations
distinct_station = distinct(transit_df, line, station_name) %>% 
  nrow() %>% 
  print()

# total distinct stations that are ADA compliant
ada_compliant = filter(transit_df, ada == TRUE) %>% 
  distinct(line, station_name) %>% 
  nrow() %>% 
  print()

# proportion of station entrances / exits without vending allow entrance?
    # no vending / total entrances
entry_no_vending = filter(transit_df, vending == "NO", entry = TRUE) %>% 
  nrow() %>% 
  print()

total_entry = filter(transit_df, entry = TRUE) %>% 
  nrow() %>% 
  print()

proportion_nvending_yentry = (entry_no_vending / total_entry) * 100
proportion_nvending_yentry

#Reformatting data
reformatted_transit_df = transit_df %>% 
  mutate(route1 = as.character(route1), 
         route2 = as.character(route2), 
         route3 = as.character(route3), 
         route4 = as.character(route4), 
         route5 = as.character(route5), 
         route6 = as.character(route6), 
         route7 = as.character(route7), 
         route8 = as.character(route8), 
         route9 = as.character(route9), 
         route10 = as.character(route10), 
         route11 = as.character(route11)
  ) %>% 
  pivot_longer(
  route1 : route11,
  names_to = "route_number",
  values_to = "train_route"
  )

# total distinct stations that serve train A
distinct_serving_a_station = filter(reformatted_transit_df, train_route == "A") %>% 
  distinct(line, station_name) %>% 
  nrow() %>% 
  print()

# total distinct stations that serve train A and is ADA compliant
distinct_serving_a_station_ada = filter(reformatted_transit_df, train_route == "A", ada == TRUE) %>% 
  distinct(line, station_name) %>% 
  nrow() %>% 
  print()
```
After cleaning the data set, there is now a total of `r nrow(transit_df)` rows and `r ncol(transit_df)` columns. The dataset contains 19 columns which includes the train line, station name, latitue and longitude of the train station, 11 train routes, if an entry exists, if there is vending, type of entrance and if the train station is ADA compliant. I cleaned the data set by dropping unneeded columns, and changing data types for better analysis. The data set is not tidy as it could be tider; for example, routes 1 through 11 could be tidier using pivot_long. 

There are a total of `r distinct_station` distinct stations.

There are a total of `r ada_compliant` ADA compliant stations.

`r proportion_nvending_yentry`% of station entrances / exits without vending allow entrance.

There are a total of `r distinct_serving_a_station` distinct stations that serve the A train.

Of the distinct stations that serve the A train, `r distinct_serving_a_station_ada` are ADA compliant.

### Question 2
```{r Q2}
mr_trash_df = read_excel("./data/202409 Trash Wheel Collection Data.xlsx", 
  sheet = "Mr. Trash Wheel", 
  range = "A2:N653") %>% #omit non-data entries
  janitor::clean_names() %>% #reasonable variable names
  filter(dumpster == "") %>% #omit rows that do not include dumpster-specific data
  mutate(
    sports_balls = as.integer(sports_balls),
    wheel = "mr_trash"
  )

prof_trash_df = read_excel("./data/202409 Trash Wheel Collection Data.xlsx", 
  sheet = "Professor Trash Wheel", 
  range = "A2:M120") %>% #omit non-data entries
  janitor::clean_names() %>% #reasonable variable names
  filter(dumpster == "") %>% #omit rows that do not include dumpster-specific data
  mutate(
    wheel = "prof_trash"
  )

gwynnda_trash_df = read_excel("./data/202409 Trash Wheel Collection Data.xlsx", 
  sheet = "Gwynnda Trash Wheel", 
  range = "A2:L265") %>% #omit non-data entries
  janitor::clean_names() %>% #reasonable variable names
  filter(dumpster == "") %>% #omit rows that do not include dumpster-specific data
  mutate(
    wheel = "gwynnda_trash"
  )
  

```

### Question 3
```{r Q3}
#Importing data
bakers_df = read_csv("./data/gbb_datasets/bakers.csv")%>% 
  janitor::clean_names() %>% 
  separate(
    baker_name, into = c("baker", "baker_last_name"), sep = " "
  ) %>% 
  arrange(series)
bakers_df

bakes_df = read_csv("./data/gbb_datasets/bakes.csv")%>% 
  janitor::clean_names()
bakes_df

results_df = read_csv("./data/gbb_datasets/results.csv", skip = 2) %>% 
  janitor::clean_names() 
results_df

#completeness / correctness across datasets
bakers_anti = anti_join(bakers_df, bakes_df, by = c("baker", "series")) #25 bakers discrepancy between bakers_df and bakes_df
resulting_anti = anti_join(bakers_df, results_df) # 1 baker discrepancy between bakers_df and results_df -- Jo Wheatley with series 2 and age 41
  #Looking over the results table, we see there is a Joanne that makes that information --> Jo most likely is Joanne

 #Changes Jo to Joanne
bakers_df = mutate(bakers_df, baker = case_match(
  baker, "Jo" ~ "Joanne")
  )

bakes_df = mutate(bakes_df, baker = case_match(
  baker, "Jo" ~ "Joanne")
  )

results_df = mutate(results_df, baker = case_match(
  baker, "Jo" ~ "Joanne")
  )

#Merging bakers with bakes first
bakers_bakes = left_join(bakers_df, bakes_df, by = c("series", "baker"))

#Merging bakers and bakes with results next
baking_merge = left_join(bakers_df, results_df, by = c("series", "baker"))


```